---
title: "Scenario1"
author: "Saim Zafar"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Import Libraries
library(stats)
library(readr)
library(astsa)
library(rugarch)
library(MTS)
library(ggplot2)
library(fGarch)
library(forecast)
library(gridExtra)
library(tidyverse)
library(tseries)
library(readr)
library(urca)
library(MASS)
```

### Scenario 1: Hydrological time series

Import the Data and plot it to see the time series.

```{r}
df_txt <- read.table(file = "~/Desktop/Stat 443/Final Report/STAT 443 - Spring 2024 Data/hydrology_data.txt", header = TRUE)
hydro_TS <- ts(df_txt)
#View(df_txt)
#head(df_txt)
#plot(df_txt$index, df_txt$Value, type = "l", xlab = "Time", ylab = "Water Levels", main = "Hydrological time series")
plot(hydro_TS,  xlab = "Time", ylab = "Water Levels", main = "Hydrological time series")
```

We should check this data for stationarity first and then see if the ACF and PACF can suggest a good model.

```{r,warning=FALSE}
kpss.test(hydro_TS)
```
We see that as the p-value is smaller than printed p-value so we have to reject the hypothesis that the data is stationary.
```{r}
par(mfrow = c(1,2))
acf(hydro_TS, lag.max = 20) 
pacf(hydro_TS, lag.max = 20)
par(mfrow = c(1,1))
```
The ACF and PACF suggest we should use Auto.ARIMA to find appropriate orders of the ARIMA model, as we don't see any simple AR or MA models in the data..
```{r}
Model <- auto.arima(hydro_TS)
summary(Model)
```
We get a ARIMA(5,1,1) model. After seeing the residuals, I am not sure if their is seasonality or not. So I will use both models as well as ETS and then use cross-validation to find the best models.  

```{r}
mod1 <- sarima(hydro_TS,p = 5,d = 1,q = 1,P = 0,Q = 1, D = 1, S = 12, details = FALSE)
mod2 <- sarima(hydro_TS,p = 5,d = 1,q = 1,P = 0,Q = 0, D = 0, S = 4,details = FALSE)
```
The diagnositics of the above models are as follows.
```{r, fig.width=5,fig.height=5}
par(mfrow = c(2,2))
tsdiag(mod1$fit)
tsdiag(mod2$fit)
par(mfrow = c(1,1))
```
We see that the residuals are similar to White Noise as suggested by the BLP test for both seasonal and non-seasonal models.

We also check a ETS model:

```{r}
ets(hydro_TS)
plot(ets(hydro_TS))
```

To compare the three models we need to perform a Cross Validation:
```{r, warning=FALSE}
n_fit <- 500
base_dat <- hydro_TS
test_id <- 1:(length(hydro_TS)-n_fit-24)
err1 <- err2 <-  err3 <- rep(NA,length(test_id))


for(i in test_id){
  #cat(paste0(i,', '))
  dat_fit <- ts(hydro_TS[1:(n_fit+i-1)])
  for1  <- as.numeric(sarima.for(dat_fit,n.ahead = 24,
                                 p = 5, d = 1, q = 1,
                                 P = 0, D = 1, Q = 1,
                                 S = 12, plot = FALSE)$pred)
  err1[i] <- sum((base_dat[(n_fit+i+0):(n_fit+i+23)] - for1)^2)
  
  for2  <- as.numeric(predict(ets(dat_fit),24)$mean)
  err2[i] <- sum((base_dat[(n_fit+i+0):(n_fit+i+23)] - for2)^2)
  
  for3  <- as.numeric(sarima.for(dat_fit,n.ahead = 24,
                                 p = 5, d = 1, q = 1,
                                 P = 0, D = 0, Q = 0,
                                 S = 4, plot = FALSE)$pred)
  err3[i] <- sum((base_dat[(n_fit+i+0):(n_fit+i+23)] - for3)^2)
}

mean(err1);mean(err2);mean(err3)
```
After CV we see that the $SARIMA(5,1,1,0,1,1)_{12}$ model is the best. Lets setup forecast into it for the next 24 months and export it to a txt file.
```{r}
forcast_Scenario1 <- sarima.for(hydro_TS, n.ahead = 24, p = 5, d = 1, q = 1,
                                 P = 0, D = 1, Q = 1,
                                 S = 12, plot.all = FALSE, plot = FALSE)

forecasts_plot <- data.frame('x'=seq(577:600),
             'forecast'=as.numeric(forcast_Scenario1$pred),
             'low'=as.numeric(forcast_Scenario1$pred-1.96*forcast_Scenario1$se),
             'upp'=as.numeric(forcast_Scenario1$pred+1.96*forcast_Scenario1$se) )
#forecasts_plot


ggplot() +
  geom_line(aes(x=seq(1:576), 
                y=hydro_TS),linewidth=1) +
  geom_line(aes(x=576+seq(1:24), 
                y=forecasts_plot$forecast),linewidth=1,color='red') +
  geom_ribbon(aes(x=576+seq(1:24),ymin=forecasts_plot$low,ymax=forecasts_plot$upp),alpha=0.25,fill='red')+
  xlab("Time")+ylab("Water Levels")
```
```{r, include= FALSE}
last.name = "Zafar"
student.id = 20888160

forecast1 <- forecasts_plot$forecast  # Forecast is a vector length 24 for forecast 1
write(forecast1, file = paste("Scenario1_",last.name,student.id,".txt", sep = ""), ncolumns = 1 )
```


